A Phenomenological Model for the Intra-Cluster Medium that matches X-ray and Sunyaev-Zel'dovich observations

### 2013-09-20 paper1.tex
Last Fabio Version

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  ...
  We build a mock galaxy cluster catalog from the large MultiDark $N$-body $\Lambda$CDM 
  simulation by adopting a phenomenological gas density model for each cluster based on X-ray
  measurements that matches Sunyaev-Zel'dovich and X-ray scaling relations and
  luminosity function. 
  ...
  We make our cosmologically complete multi-frequency mock catalogs for 
  the (non-)thermal cluster emission at different redshifts publicly and freely available on-line
  through the MultiDark database.
\end{abstract}

@@

\begingroup
\let\thefootnote\relax\footnotetext{* f.zandanel@uva.nl}
\let\thefootnote\relax\footnotetext{$\dagger$ christoph.pfrommer@h-its.org}
\let\thefootnote\relax\footnotetext{$\S$ fprada@iaa.es}
\endgroup

{\bf Introduction...}

%
%dark matter (DM)  
%

{\bf Our original motivation to build such a mock catalog is to predict the 
non-thermal emission at radio and gamma-ray frequencies of a cosmological 
complete sample of galaxy clusters. This is presented in a companion 
paper (REF, hereafter Paper~2).
The two basic ingredients of the two works are an $N$-body, \emph{DM-only}
cosmological simulation that we use in constructing our complete cluster sample,
and the (non-)thermal emission model. Here, we use the MultiDark simulation that
will be described in Section~\ref{sec:2}, along with our final cluster sample. As described 
in Paper~2, for any given cluster, the necessary ingredients for the non-thermal emission 
model are mass, temperature, and gas density distribution. We assign to
each cluster in that simulation a gas density profile that is phenomenologically
constructed from state-of-art X-ray observations, as will be shown in
Section~\ref{sec:3}. In Section~4 and 5, we will show that this approach can successfully
reproduce the known X-ray cluster characteristics, such as the X-ray luminosity
function (XLF), the luminosity-mass relation, $L_{\rmn{X}}- M$, and the
$Y_{\rmn{X}}-M$ relation, where $Y_{\rmn{X}}=M_\rmn{gas}k_{\rmn{B}}T$ with an
X-ray-derived gas mass $M_{\rmn{gas}}$ and temperature $T$
\citep{2006ApJ...650..128K}. Moreover, we compare our $Y_{\rmn{SZ}}-M$ relation
with SZ-derived measurements. Only if our model matches available cluster data
on the gas properties, we can safely use it in Paper~2.} In this work, the cluster mass $M_{\Delta}$ 
and radius $R_{\Delta}$ are defined with respect to a density that is $\Delta=200$ or 
$\Delta=500$ times the \emph{critical} density of the Universe. We adopt density 
parameters of $\Omega_{\rmn{m}}=0.27$, $\Omega_{\Lambda}=0.73$ and today's 
Hubble constant of $H_0 = 100 \times h_{70}$~km~s$^{-1}$~Mpc$^{-1}$ where $h_{70} = 0.7$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{MultiDark simulation and final cluster sample}
\label{sec:2}
The MultiDark simulation used in this work is described in detail in \cite{2011arXiv1104.5130P}.  
It is an $N$-body cosmological simulation done with the Adaptive-Refinement-Tree (ART) 
code \citep{1997ApJS..111...73K} of $2048^3$ particles within a ($1000$~Mpc~$h^{-1}$)$^3$ 
cube. The adopted cosmological parameters are $\Omega_{\rmn{m}}=0.27$, $\Omega_{\Lambda}=0.73$,
$\Omega_\rmn{b}=0.0469$, $n_\rmn{s}=0.95$, $h=0.7$ and $\sigma_8=0.82$. This simulation is 
particularly well suited for our purpose because of its large 
number of high-mass objects, i.e., galaxy clusters.

We use the MultiDark halo catalog from its on-line database,\footnote{www.multidark.org.} 
constructed with the Bound Density Maxima algorithm \citep{1997astro.ph.12217K}.  
We will mainly use $M_{500}$ and $R_{500}$ for comparison with existing observational works.  
We use the technique described in \cite{2003ApJ...584..702H} to convert $M_{200}$ and
$R_{200}$ provided by the MultiDark halo catalog to $M_{500}$ and $R_{500}$.  In
creating our cluster sample we only select distinct halos, i.e., those halos that
are not sub-halos of any other halo, which by definition are not galaxy clusters.

@@

{\bf In Paper~2, we present predictions for the LOFAR radio observatory
which is expected to detect diffuse radio emission in clusters up to redshift $z \approx 1$ (e.g.,
\citealp{2012JApA..tmp...34R}).} 
Thus, we make use of different simulation snapshots up to $z=1$. 
{\bf The extrapolation of our model beyond this redshift is rather uncertain as it is based
on observations at low(er) redshift.}
In Table~\ref{tab:z}, we show the total cluster number in our final cluster sample at different redshifts.

@@

We decided to use a \emph{phenomenological} approach and to construct the gas
density profiles directly from X-ray observations. A suitable X-ray sample that
provides the needed information is the \emph{Representative XMM-Newton Cluster
  Structure Survey} (REXCESS) sample \citep{2008A&A...487..431C,
  2009A&A...498..361P}. It is a sample of 31 galaxy clusters of different
dynamical states at redshifts $0.06<z<0.18$ with detailed information on the
de-projected electron density profiles \citep{2008A&A...487..431C}. In
Fig.~\ref{fig:gas_profiles}, we show the 31 electron density profiles of the
REXCESS sample color-coded by CCCs and NCCCs.

@@

\begin{figure*} 
\centering
\includegraphics[width=0.48\textwidth]{figures/lx_m.eps}
\includegraphics[width=0.48\textwidth]{figures/yx_m.eps}
%\includegraphics[width=0.33\textwidth]{figures/sz_m.eps}
\caption{X-ray scaling relations. Grey triangles show the MultiDark
  sample (limited to the mass range covered by observations), the black line is
  the corresponding scaling relation, and the blue line is the
  observational result. The black crosses represent the median values of the
  quantity in question for a given mass bin (indicated by horizontal error
  bars), and the vertical error bars represent the standard deviation within a
  bin.  \emph{Left.} We compare the bolometric X-ray luminosity-to-mass
  relation, $L_{\rmn{bol}}-M_{500}$, at $z=0.2$ to the observational sample by
  \protect\cite{2010MNRAS.406.1773M} with a median of $z \approx 0.2$. 
  \emph{Right.} $Y_{\rmn{X}}-M_{500}$ scaling relation of our model in comparison to the
  observational sample by \protect\cite{2010MNRAS.406.1773M}.} 
  %\emph{Right.} $Y_{\rmn{SZ}}-M_{500}$ scaling relation at $z=0.2$ in comparison
  %to the observational sample by the \protect\cite{2013arXiv1303.5080P} with a 
  %median redshift of about $0.18$. The bottom panels show the relative difference to
  %the observational scaling relations.}
\label{fig:X_LM}
\end{figure*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{SZ scaling relations}
\label{sec:5}
In the right panel of Fig.~\ref{fig:X_LM}, we compare the
$Y_{\rmn{SZ}}-M_{500}$ relation in our model (calculated as in equation~(3) of
\citealp{2011arXiv1109.3709B}) with the observed scaling relation by the
\cite{2013arXiv1303.5080P}. {\bf We compare with the result obtained from the 
\emph{Planck} COSMO sample with Malmquist bias correction (see their Table~A.1),
which has a median redshift of $0.18$. Hence, we compare the data to our relation 
at the MultiDark snapshot $z=0.2$}. Our model reproduces the data remarkably well, except 
for the high-mass end where our simulations have a weaker constraining power due to the 
smaller box size in comparison to the survey volume of {\em Planck}. We find a scatter of
$\sigma_{yx} \approx 0.11$ which compares well with the \emph{Planck} result of
$\sigma_{yx} \approx 0.08$. In Table~\ref{tab:YSZfits}, we report our SZ scaling
relations for different redshifts.

\begin{figure} 
\centering
\includegraphics[width=0.48\textwidth]{figures/sz_m.eps}
\caption{SZ scaling relations. Grey triangles show the MultiDark
  sample (limited to the mass range covered by observations), the black line is
  the corresponding scaling relation, and the blue line is the
  observational result. The black crosses represent the median values of the
  quantity in question for a given mass bin (indicated by horizontal error
  bars), and the vertical error bars represent the standard deviation within a
  bin.  \emph{Left.} We compare the bolometric X-ray luminosity-to-mass
  relation, $L_{\rmn{bol}}-M_{500}$, at $z=0.2$ to the observational sample by
  \protect\cite{2010MNRAS.406.1773M} with a median of $z \approx 0.2$. 
  \emph{Right.} $Y_{\rmn{X}}-M_{500}$ scaling relation of our model in comparison to the
  observational sample by \protect\cite{2010MNRAS.406.1773M}. 
  \emph{Left.} $Y_{\rmn{SZ}}-M_{500}$ scaling relation at $z=0.2$ in comparison
  to the observational sample by the \protect\cite{2013arXiv1303.5080P} with a 
  median redshift of about $0.18$. The bottom panels show the relative difference to
  the observational scaling relations.}
\label{fig:SZ_M}
\end{figure}

\begin{figure*} 
\centering
\includegraphics[width=0.48\textwidth]{figures/sz_m_SPT.eps}
\includegraphics[width=0.48\textwidth]{figures/sz_m_ACT.eps}
\caption{SZ scaling relations. Grey triangles show the MultiDark
  sample (limited to the mass range covered by observations), the black line is
  the corresponding scaling relation, and the blue line is the
  observational result. The black crosses represent the median values of the
  quantity in question for a given mass bin (indicated by horizontal error
  bars), and the vertical error bars represent the standard deviation within a
  bin.  \emph{Left.} We compare the bolometric X-ray luminosity-to-mass
  relation, $L_{\rmn{bol}}-M_{500}$, at $z=0.2$ to the observational sample by
  \protect\cite{2010MNRAS.406.1773M} with a median of $z \approx 0.2$. 
  \emph{Right.} $Y_{\rmn{X}}-M_{500}$ scaling relation of our model in comparison to the
  observational sample by \protect\cite{2010MNRAS.406.1773M}. 
  \emph{Left.} $Y_{\rmn{SZ}}-M_{500}$ scaling relation at $z=0.2$ in comparison
  to the observational sample by the \protect\cite{2013arXiv1303.5080P} with a 
  median redshift of about $0.18$. The bottom panels show the relative difference to
  the observational scaling relations.}
\label{fig:SZ_M_1}
\end{figure*}

@@

{\bf We additionally compare our $Y_{\rmn{SZ}}$ to the results from the South Pole Telescope (SPT) 
and Atacama Cosmology Telescope (ACT) surveys.
We select a subsample of SPT clusters from \cite{2013ApJ...763..127R}, formed by the
115 objects with confirmed redshift and signal-to-noise above 5, which has a median
redshift of 0.56 with rms of 0.27. \cite{2013ApJ...763..127R} provide $Y_{\rmn{SZ}}$ 
within an aperture of $1'$, corresponding to a median radius of $0.2$~h$_{70}^{-1}$~Mpc 
at $z = 0.56$, which we use as aperture to calculate $Y_{\rmn{SZ, SPT}}$ in our MultiDark sample.

In the left panel of Fig.~\ref{fig:SZ_M_1}, we compare the SPT cluster subsample to our 
scaling relation at the three different MutliDark snapshots of $z = 0.4$, $0.61$ 
and $0.78$. We color-coded the SPT clusters in order to show that the SPT 
selection function is nearly redshift-independent. The fact that the angular size of galaxy clusters 
decreases with redshift is an advantage for SPT due to its small beam, in fact, higher-redshift clusters 
are less confused by primary CMB fluctuations. This effect is inverse for \emph{Planck} due to its
larger beam. We can conclude that while our $Y_{\rmn{SZ}}-M_{500}$ scaling relation 
compare fairly well to the median SPT cluster population, the SPT sample has a too
large scatter to be well represented by single MultiDark snapshots.

In the right panel of Fig.~\ref{fig:SZ_M_1}, we compare our scaling relation to
the ACT sample overlapping with the Sloan Digital Sky Survey stripe 82 
(\citealp{2013JCAP...07..008H}; we use the quantities inferred from the
universal pressure profile scaling relation).
The sample has a median redshift of 0.58 with rms of 0.28.
\cite{2013JCAP...07..008H} provides $Y_{\rmn{SZ}}$ within $R_{500}$ which
median, $0.42$~h$_{70}^{-1}$~Mpc, is, however, almost half of ours at the
MultiDark snapshot $z=0.61$. Therefore, we use an aperture of $0.42$~h$_{70}^{-1}$~Mpc
to calculate our $Y_{\rmn{SZ, ACT}}$. Also in this case we can conclude
that our mock sample reproduce fairly well the median observational sample. 
Note, however, the very different scatter of SPT and ACT samples.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
\label{sec:6}

{\bf Discussion...}

{\bf We build a complete cosmological sample of galaxy clusters from the MultiDark 
$N$-body simulation with redshifts ranging from $z = 0$ to 1. We construct a 
\emph{phenomenological} model for the cluster-gas distribution. This is characterized 
by X-ray-inferred (CCC and NCCC) gas profiles (taken from the REXCESS sample) 
and a cluster-mass-dependent gas fraction. We assign such a (cluster mass-dependent) 
gas density profile to each DM halo in our sample and sort it into the NCCC/CCC populations 
according to a dynamical disturbance parameter that is calculated from the DM distribution.
With this model, we obtain a cosmologically complete mock catalog of galaxy
clusters that matches the observed $L_{\rmn{X, bol}}$-to-$M_{500}$,
$Y_{\rmn{X}}$-to-$M_{500}$, and $Y_{\rmn{SZ}}$-to-$M_{500}$ relations, as well
as the X-ray luminosity function. We apply it in the companion Paper~2 to 
predict the radio and gamma-ray emission of our sample.

Eventually, we provide a cosmologically complete multi-frequency mock catalog for the (non-)thermal 
cluster emission at different redshifts. We make these catalogs publicly and freely available 
on-line through the MultiDark database (www.multidark.org). Those contain the quantities
$\rho_{\rmn{gas}}$, $L_{\rmn{X, bol}}$, $Y_{\rmn{X}}$, $Y_{\rmn{SZ}}$, $L_{120~\rmn{MHz}}$, 
$L_{1.4~\rmn{GHz}}$, $L_{>100~\rmn{MeV}}$ and $L_{>100~\rmn{GeV}}$ (among many others). 
We hope that the community can make valuable use of these catalogues in synergy with the future 
radio, X-ray and gamma-ray data.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgments}
We thank Anders Pinzke for many useful discussions. We also thank Harald Ebeling, 
Stefan Gottl{\"o}berg, Anatoly Klypin, Andrey Kravtsov, Adam Mantz, Frazer Pearce, 
Jos\'e Alberto Rubi\~no, and Gustavo Yepes, for the useful discussions and advices. 
Finally, we thank the MultiDark database people, in particular Adrian Partl and Kristin Riebe. 
F.Z.{\ }acknowledges the CSIC financial support as a JAE-Predoc grant of the 
program ``Junta para la Ampliaci\'on de Estudios'' co-financed by the FSE. 
F.Z.{\ }acknowledges the hospitality of the Leiden Observatory during his stay.
F.Z.{\ }and F.P.{\ } thank the support of the Spanish MICINN's Consolider-Ingenio 2010 Programme under
grant MultiDark CSD2009-00064. C.P.{\ }gratefully acknowledges financial support
of the Klaus Tschira Foundation. The MultiDark Database used in this paper and
the web application providing online access to it were constructed as part of
the activities of the German Astrophysical Virtual Observatory as result of a
collaboration between the Leibniz-Institute for Astrophysics Potsdam (AIP) and
the Spanish MultiDark Consolider Project CSD2009-00064. The Bolshoi and
MultiDark simulations were run on the NASA's Pleiades supercomputer at the NASA
Ames Research Center.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  {\bf We construct a phenomenological gas density model for the intra-cluster medium
  based on X-ray inferred densities for cool-core and non cool-core clusters, and a 
  cluster-mass-dependent gas fraction. We then build a mock galaxy cluster 
  catalog from the large MultiDark $N$-body $\Lambda$CDM simulation by assigning
  such a (mass-dependent) gas density profile to each dark matter halo in our sample, 
  which is sorted into cool-core and non cool-core clusters according to a dynamical 
  disturbance parameter. Our gas density model permits to match the observed 
  Sunyaev-Zel'dovich and X-ray scaling relations and luminosity function. Therefore,
  it can be reliable applied to make predictions for present and future instruments. 
  We apply our catalogs to calculate non-thermal radio and gamma-ray emission of galaxy 
  clusters in a companion paper. Eventually, we make our cosmologically complete multi-frequency 
  mock catalogs for the (non-)thermal cluster emission at different redshifts publicly and freely 
  available on-line through the MultiDark database.}
\end{abstract}

@@

\begingroup
\let\thefootnote\relax\footnotetext{* f.zandanel@uva.nl}
\let\thefootnote\relax\footnotetext{$\dagger$ christoph.pfrommer@h-its.org}
\let\thefootnote\relax\footnotetext{$\S$ fprada@iaa.es\\}
\endgroup

{\bf The study of the (non-)thermal emission in galaxy clusters provides
important information on structure formation process and on 
fundamental characteristics of our Universe. This field is at a 
crucial point of its development thanks to the very low sensitivity of present and future-generation 
instruments such as LOFAR\footnote{www.lofar.org}, \emph{e}ROSITA\footnote{www.mpe.mpg.de/eROSITA}, 
ASTRO-H\footnote{astro-h.isas.jaxa.jp} and CTA\footnote{www.cta-observatory.org}, to name but a few. 
Therefore, it is important to provide prospects for future surveys and
to understand in which way they can improve our current knowledge.
To this end, a complete cosmological mock sample of galaxy
clusters is needed. This can be realized either using analytical 
expressions for the cluster mass function (e.g., \citealp{2001MNRAS.321..372J}) or cosmological cluster 
simulations. In this work, we make use of the recent MultiDark $N$-body
simulation (\citealp{2011arXiv1104.5130P}; www.multidark.org).

Our main motivation to build such a mock catalog has been to predict the 
non-thermal emission at radio frequencies of a cosmological 
complete sample of galaxy clusters. This is presented in a companion 
paper (\citealp{paper2}; hereafter Paper~II). As described in Paper~II, for 
any given cluster, fundamental ingredients for the non-thermal emission 
model are the mass and gas density distribution. The MultiDark 
simulation used here is a $N$-body, \emph{dark matter (DM)-only} simulation and, 
therefore, does not contain any information on the intra-cluster medium (ICM) gas  
distribution. A cosmological complete sample of clusters can 
only be obtained from $N$-body simulations as hydrodynamical
simulations of such large volumes are computationally too expensive.
Therefore, we need a method to determine the ICM properties
of any given cluster from its total mass.

The simplest approach is to assume that clusters are self-similar with 
the thermal properties determined by their mass \citep{1986MNRAS.222..323K}.
However, X-ray observations showed that clusters are extremely 
complex systems. In fact, we know that self-similarity is broken, especially at 
low-mass scales (see \citealp{2005RvMP...77..207V} for a review).
The study of the thermodynamical condition of clusters is crucial 
to understand how self-similarity is broken. One possible solution 
is, for example, energy feedback from active galactic nuclei (e.g.,
\citealp{2013arXiv1303.5443P}).

In order to build a cluster sample with realistic ICM densities, we 
adopt a \emph{phenomenological} approach and construct our density 
profiles from X-ray observations. We then compare the characteristics of 
the clusters in our mock sample to observational results. We study the X-ray 
luminosity function (XLF), the luminosity-mass relation, $L_{\rmn{X}}- M$, and 
the $Y_{\rmn{X}}-M$ relation, where $Y_{\rmn{X}}=M_\rmn{gas}k_{\rmn{B}}T$ 
with an X-ray-derived gas mass $M_{\rmn{gas}}$ and temperature $T$
\citep{2006ApJ...650..128K}. Moreover, we consider the thermal 
Sunyaev-Zel’dovich (SZ) effect and compare our $Y_{\rmn{SZ}}-M$ 
relation with recent SZ-derived measurements. Only if our model matches
available cluster data on the gas properties, we can safely use it in Paper~II
and be sure that we will obtain realistic predictions.
The final result is a complete cosmological cluster mock catalog, at
different redshifts, containing information on the DM and ICM properties 
of each object, together with its (non-)thermal emission at different 
frequencies. This mock catalog reproduces existing observations
and is useful in making predictions for future surveys.

In next Section~\ref{sec:2}, we introduce the MultiDark simulation
along with our final cluster sample. We assign to each of our clusters  
the phenomenologically constructed gas density profile in 
Section~\ref{sec:3}. In Sections~\ref{sec:4} and 
\ref{sec:5}, we show that this approach can successfully reproduce 
the known X-ray and SZ cluster characteristics. We present the
resulting mock cluster catalogs in Section~\ref{sec:6}, and eventually 
summarize our conclusions in Section~\ref{sec:7}.
In this work, the cluster mass $M_{\Delta}$ and radius $R_{\Delta}$ are defined 
with respect to a density that is $\Delta=200$ or $\Delta=500$ times the \emph{critical} 
density of the Universe. We adopt density parameters of $\Omega_{\rmn{m}}=0.3$, 
$\Omega_{\Lambda}=0.7$ and today's Hubble constant of 
$H_0 = 100 \times h_{70}$~km~s$^{-1}$~Mpc$^{-1}$ where $h_{70} = 0.7$.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{MultiDark simulation and final cluster sample}
\label{sec:2}
The MultiDark simulation used in this work is described in detail in \cite{2011arXiv1104.5130P}.  
It is an $N$-body cosmological simulation done with the Adaptive-Refinement-Tree (ART) 
code \citep{1997ApJS..111...73K} of $2048^3$ particles within a ($1000$~Mpc~$h^{-1}$)$^3$ 
cube. {\bf The adopted cosmological parameters are $\Omega_{\rmn{m}}=0.27$, $\Omega_{\Lambda}=0.73$,
$\Omega_\rmn{b}=0.0469$, $n_\rmn{s}=0.95$, $h=0.7$ and $\sigma_8=0.82$.} This simulation is 
particularly well suited for our purpose because of its large 
number of high-mass objects, i.e., galaxy clusters.

We use the MultiDark halo catalog from its on-line database,\footnote{www.multidark.org} 
constructed with the Bound Density Maxima (BDM) algorithm \citep{1997astro.ph.12217K}.  
We will mainly use $M_{500}$ and $R_{500}$ for comparison with existing observational works.  
We use the technique described in \cite{2003ApJ...584..702H} to convert $M_{200}$ and
$R_{200}$ provided by the MultiDark halo catalog to $M_{500}$ and $R_{500}$.  In
creating our cluster sample we only select distinct halos, i.e., those halos that
are not sub-halos of any other halo, which by definition are not galaxy clusters.

@@

{\bf In Paper~II, we present predictions for the LOFAR radio observatory
which is expected to detect diffuse radio emission in clusters up to redshift $z \approx 1$ (e.g.,
\citealp{2012JApA..tmp...34R}).} Thus, we make use of different simulation snapshots up to $z=1$. 
{\bf The extrapolation of our model beyond this redshift is rather uncertain as it is based
on observations at low(er) redshift.} In Table~\ref{tab:z}, we show the total cluster number in 
our final cluster sample at different redshifts.

@@

We decided to use a \emph{phenomenological} approach and to construct the gas
density profiles directly from X-ray observations. A suitable X-ray sample that
provides the needed information is the \emph{Representative XMM-Newton Cluster
  Structure Survey} (REXCESS) sample \citep{2008A&A...487..431C,
  2009A&A...498..361P}. It is a sample of 31 galaxy clusters of different
dynamical states at redshifts $0.06<z<0.18$ with detailed information on the
de-projected electron density profiles \citep{2008A&A...487..431C}. In
Fig.~\ref{fig:gas_profiles}, we show the 31 electron density profiles of the
REXCESS sample color-coded by non cool-core clusters (NCCCs) and 
cool-core clusters (CCCs).

@@

\begin{figure*} 
\centering
\includegraphics[width=0.48\textwidth]{figures/lx_m.eps}
\includegraphics[width=0.48\textwidth]{figures/yx_m.eps}
%\includegraphics[width=0.33\textwidth]{figures/sz_m.eps}
\caption{{\bf X-ray scaling relations. Grey triangles show the MultiDark
  sample (limited to the mass range covered by observations), the black line is
  the corresponding scaling relation, and the blue line is the
  observational result. The black crosses represent the median values of the
  quantity in question for a given mass bin (indicated by horizontal error
  bars), and the vertical error bars represent the standard deviation within a
  bin.  \emph{Left.} We compare the bolometric X-ray luminosity-to-mass
  relation, $L_{\rmn{bol}}-M_{500}$, at $z=0.2$ to the observational sample by
  \protect\cite{2010MNRAS.406.1773M} with a median of $z \approx 0.2$. 
  \emph{Right.} $Y_{\rmn{X}}-M_{500}$ scaling relation of our model in comparison to the
  observational sample by \protect\cite{2010MNRAS.406.1773M}.}} 
  %\emph{Right.} $Y_{\rmn{SZ}}-M_{500}$ scaling relation at $z=0.2$ in comparison
  %to the observational sample by the \protect\cite{2013arXiv1303.5080P} with a 
  %median redshift of about $0.18$. The bottom panels show the relative difference to
  %the observational scaling relations.}
\label{fig:X_LM}
\end{figure*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{SZ scaling relations}
\label{sec:5}
In the right panel of Fig.~\ref{fig:X_LM}, we compare the
$Y_{\rmn{SZ}}-M_{500}$ relation in our model (calculated as in equation~(3) of
\citealp{2011arXiv1109.3709B}) with the observed scaling relation by the
\cite{2013arXiv1303.5080P}. {\bf We compare with the result obtained from the 
\emph{Planck} COSMO sample with Malmquist bias correction,
which has a median redshift of $0.18$. Hence, we compare the data to our relation 
at the MultiDark snapshot $z=0.2$}. Our model reproduces the data remarkably well, except 
for the high-mass end where our simulations have a weaker constraining power due to the 
smaller box size in comparison to the survey volume of {\em Planck}. We find a scatter of
$\sigma_{yx} \approx 0.11$ which compares well with the \emph{Planck} result of
$\sigma_{yx} \approx 0.08$. In Table~\ref{tab:YSZfits}, we report our SZ scaling
relations for different redshifts.

\begin{figure} 
\centering
\includegraphics[width=0.48\textwidth]{figures/sz_m.eps}
\caption{{\bf SZ scaling relation of our sample at $z=0.2$ in comparison to the observational 
  sample by \protect\cite{2013arXiv1303.5080P} with a median redshift of about $0.18$.
  Grey triangles show the MultiDark sample (limited to the mass range covered by observations), 
  the black line is the corresponding scaling relation, and the blue line is the observational result. 
  The black crosses represent the $Y_{\rmn{SZ}}$ median values of our sample for a given mass bin (indicated by 
  horizontal error bars), and the vertical error bars represent the standard deviation within a
  bin. The bottom panels show the relative difference to the observational scaling relations.}}
\label{fig:SZ_M}
\end{figure}

\begin{figure*} 
\centering
\includegraphics[width=0.48\textwidth]{figures/sz_m_SPT.eps}
\includegraphics[width=0.48\textwidth]{figures/sz_m_ACT.eps}
\caption{{\bf SZ scaling relation of our sample at $z=0.61$ compared with the SPT and ACT results. 
  Triangles show the MultiDark sample (limited to the mass range covered by observations) 
  while the crosses represent the observational result. \emph{Left.} Comparison with the 
  SPT sample \protect\citep{2013ApJ...763..127R} color-coded by redshift. 
  $Y_{\rmn{SZ, SPT}}$ is given within an aperture of  $0.2$~h$_{70}^{-1}$~Mpc.
  \emph{Right.} Comparison with the ACT sample \protect\citep{2013JCAP...07..008H} color-coded 
  by redshift. $Y_{\rmn{SZ, ACT}}$ is given within an aperture of $0.42$~h$_{70}^{-1}$~Mpc.}}
\label{fig:SZ_M_1}
\end{figure*}

@@

{\bf We additionally compare our $Y_{\rmn{SZ}}$ to the results from the South Pole 
Telescope (SPT) and Atacama Cosmology Telescope (ACT) surveys.
We select a subsample of SPT clusters from \cite{2013ApJ...763..127R} by taking
only objects with confirmed redshift and signal-to-noise $\ge 5$. The resulting sample 
consists of 115 clusters with a median redshift of 0.56 and rms 0.27. Note the much larger 
scatter in redshift of the SPT sample with respect to the \emph{Planck} one that has 
a rms of 0.11. \cite{2013ApJ...763..127R} provide $Y_{\rmn{SZ}}$ 
within an aperture of $1'$, corresponding to a median radius of $0.2$~h$_{70}^{-1}$~Mpc 
at $z = 0.56$, which we use as aperture to calculate $Y_{\rmn{SZ, SPT}}$ in our MultiDark sample.

In the left panel of Fig.~\ref{fig:SZ_M_1}, we compare the SPT cluster sample to our 
scaling relation at $z = 0.61$. We color-coded the SPT clusters in order to show that the SPT 
selection function is nearly redshift-independent. The fact that the angular size of galaxy clusters 
decreases with redshift is an advantage for SPT due to its small beam, in fact, higher-redshift clusters 
are less confused by primary CMB fluctuations \citep{2013ApJ...763..127R}. This effect is inverse 
for \emph{Planck} due to its larger beam. We can conclude that while our $Y_{\rmn{SZ}}-M_{500}$ 
scaling relation compare well to the median SPT cluster population, the SPT sample has a too
large scatter to be well represented by single MultiDark snapshots.

In the right panel of Fig.~\ref{fig:SZ_M_1}, we compare our scaling relation to
the ACT sample overlapping with the Sloan Digital Sky Survey stripe 82 
\citep{2013JCAP...07..008H}. This sample has a median redshift of 0.58 with rms 
of 0.28. \cite{2013JCAP...07..008H} provides $Y_{\rmn{SZ}}$ within $R_{500}$, which
median, $0.42$~h$_{70}^{-1}$~Mpc, is almost half of ours at the
MultiDark snapshot $z=0.61$. Therefore, we use an aperture of $0.42$~h$_{70}^{-1}$~Mpc
to calculate our $Y_{\rmn{SZ, ACT}}$. 
The fact that our $R_{500}$ does not compare well with the ACT values, while it does with 
\emph{Planck}, may be explained as follows. We adopt the ACT measurements derived from 
the 148~GHz data assuming that the cluster pressure is described by the universal 
profile of \cite{2010A&A...517A..92A}. In contrast, \cite{2013arXiv1303.5080P}
derived $R_{500}$ from X-ray measurements of $M_{500}$, similarly to
our approach of constructing gas densities from an X-ray sample of clusters.

Summarizing, our SZ scaling relations compare very well to the result from
the \emph{Planck} collaboration. However, the comparison with the SPT and 
ACT samples is not as straightforward due to the different selection function 
of these telescopes with respect to \emph{Planck}, which results in a much
more (redshift) non-homogenous sample. Nevertheless, our mock clusters 
can reproduce well the median SPT and ACT populations. A more detailed
comparison would need the generation of light-cone mock catalogs,
including instrumental effects, which is beyond the scope of this 
work.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Multi-frequency Mock Catalogs}
\label{sec:6}

{\bf As a final product of this work, we construct cosmological complete mock catalogs of galaxy clusters, at the different redshifts of 
Table~\ref{tab:z}, for $M_{200}\geq1\times10^{14}$~$h^{-1}$~M$_{\odot}\approx1.4\times10^{14}$~$h_{70}^{-1}$~M$_{\odot}$.
We make these catalogs publicly available through the MultiDark database 
(www.multidark.org) where we will also post possible updates.

Our catalogs contain all the information regarding the DM properties of each object as 
given by the corresponding original MultiDark BDM halo catalogs. Additionally, we include the
following information:

\begin{itemize}
\item $M_{500}$ and $R_{500}$ calculated from $M_{200}$ and $R_{200}$ of the BDM catalogs with the \cite{2003ApJ...584..702H} method,
\item the ICM X-ray temperature assigned via the $T_{\rmn{ci}}-M_{500}$ relation by \cite{2010MNRAS.406.1773M},
\item a state flag identifying the cluster as NCCC or CCC as described in Section~\ref{sec:3},
\item the central gas density $\rho_{\rmn{gas},0}$ with which the full ICM radial profile can be calculated as in Section~\ref{sec:3},
\item the bolometric X-ray thermal bremsstrahlung luminosity $L_{\rmn{X, bol}}$ within $R_{500}$,
\item the $Y_{\rmn{X}}$ and $Y_{\rmn{SZ}}$ parameters within $R_{500}$.
\end{itemize}

In Paper~II, we present a model to calculate the possible radio and gamma-ray emission of 
clusters as result of the cosmic-ray (CR) proton-proton interactions with the ICM. We refer
to that paper for all the details. However, we want to point out that our mock catalogs include also:

\begin{itemize}
\item a state flag identifying whether a given object is a radio-loud or a radio-quiet cluster, 
\item the CR-to-thermal pressure averaged over the cluster volume within $R_{500}$,
\item the synchrotron radio luminosity, within $R_{500}$, at 120~MHz and 1.4~GHz,
\item the gamma-ray luminosity, within $R_{500}$, above 100~MeV and 100~GeV.
\end{itemize}
